NanoStringNorm Trial
========================================================

The following code is designed to import, process, and analyze NanoString data from Todd Golde's experiments in APP (Tg) and normal (WT) mice models, with and without over-expression of IL-10.

Load the required libraries.
```{r}
require("NanoStringNorm")
library("ggplot2")
```


We first import the data and create a data frame using `read.table`.
```{r}
NanoString.mRNA <- read.table("./data/AD_nanoString_data.txt", fill = TRUE, header = TRUE, 
                              as.is = TRUE)
str(NanoString.mRNA)
```

Some of the gene have duplicate names, so we add a simple tag to distinguish.
```{r}
duplicateNames <- duplicated(NanoString.mRNA$Name)
NanoString.mRNA$Name[duplicateNames] <- paste(NanoString.mRNA$Name[duplicateNames], "-2", 
                                              sep = "")
```

Next we define the housekeeping genes in the data.
```{r}
NanoString.mRNA[NanoString.mRNA$Name %in% c("Cltc", "GAPDH", "Hprt1", "Pgk1", "Tubb5"), 
                "Code.Class"] <- "Housekeeping"
```

Here we extract sample subsets, but just for the Tg data.
```{r}
annoCols <- colnames(NanoString.mRNA)[c(1:3)]
sampleIDs <- colnames(NanoString.mRNA)[-c(1:3)]
tgIL10Samples <- sampleIDs[c(1:3,7:9)]
tgCtrlSamples <- sampleIDs[c(4:6,10:12)]


tgNS.mRNA <- subset(x = NanoString.mRNA, 
                    select = colnames(NanoString.mRNA) %in% 
                      c(annoCols, tgIL10Samples, tgCtrlSamples))
```

We want to build a trait indicator vector to look at differential expression among Tg samples.
```{r}
tg.names <- names(tgNS.mRNA)[-c(1:3)]
IL10vsCTRL <- (tg.names %in% tgIL10Samples) + 1 # value 2 indicates IL10 class
trait.tg <- data.frame(
  row.names = tg.names,
  IL10vsCTRL = IL10vsCTRL)
```

We also need to define the biological replicates belonging to each experimental group.
```{r}
tg.bioReps <- rep("", times = length(tg.names))
tg.bioReps[tg.names %in% tgIL10Samples] <- "Tg.IL10"
tg.bioReps[tg.names %in% tgCtrlSamples] <- "Tg.CTRL"
```

The `norm.comp` function allows us to examine all combinations of normalization options. `CodeCount` represents the correction based on positive controls, `Background` the correction based on negative controls, and `SampleContent` the correction based on housekeeping genes. 
```{r}
tg.norm.comp.results.test <- norm.comp(
    x = tgNS.mRNA,
    replicates = tg.bioReps,
    CodeCount.methods = c("none", "sum", "geo.mean"),
    Background.methods = c("none", "mean", "mean.2sd", "max"),
    SampleContent.methods = c("none", "housekeeping.sum", "housekeeping.geo.mean",
                              "total.sum", "low.cv.geo.mean", 
                              "top.mean", "top.geo.mean"),
    OtherNorm.methods = "none",
    histogram = FALSE,
    verbose = FALSE,
    icc.method = "anova")
```

To reduce technical variation that is independent of NanoString assay variation, we would like to keep at least some form of correction based on expression of housekeeping genes. As such, we can try to run `norm.comp` again, but with fewer options for `SampleContent` methods.
```{r}
tg.norm.comp.results.test <- norm.comp(
    x = tgNS.mRNA,
    replicates = tg.bioReps,
    CodeCount.methods = c("none", "sum", "geo.mean"),
    Background.methods = c("none", "mean", "mean.2sd", "max"),
    SampleContent.methods = c("housekeeping.sum", "housekeeping.geo.mean"),
    OtherNorm.methods = "none",
    histogram = FALSE,
    verbose = FALSE,
    icc.method = "anova")
```

Within this scope, there isn't a noticeable difference in the ratio of biological (i.e., endogenous genes) to technical (i.e., positive controls) variation (based on coefficient of variation). The intra-class correlation is around 0.6 for most combinations of options, but this metric might not be that informative anyway (we have no clear reason to assume strong correlation among genes within each group).

We ultimately would like to compare groups across different experiments, so we'll now take a look at all of the data (including Tg, WT, and old mice samples).
```{r}
norm.comp.results.test <- norm.comp(
    x = NanoString.mRNA,
    CodeCount.methods = c("none", "sum", "geo.mean"),
    Background.methods = c("none", "mean", "mean.2sd", "max"),
    SampleContent.methods = c("housekeeping.sum", "housekeeping.geo.mean"),
    OtherNorm.methods = "none",
    histogram = FALSE,
    verbose = TRUE)
```

Using the geometric mean of housekeeping genes to scale endogenous genes appears to produce a higher biological to technical CV ratio, so we'll further reduce the `SampleContent` options. We can also remove the **none** options, as we definitely want to include some form of correction based on positive and negative controls.
```{r}
norm.comp.results.test <- norm.comp(
    x = NanoString.mRNA,
    CodeCount.methods = c("sum", "geo.mean"),
    Background.methods = c("mean", "mean.2sd", "max"),
    SampleContent.methods = c("housekeeping.geo.mean"),
    OtherNorm.methods = "none",
    histogram = FALSE,
    verbose = TRUE)
```

All of the above combinations of options yield virtually identical results for normalization. Given this observation, we'll go ahead and use the *recommended* set of options in the `NanoStringNorm` manual. Some of our subsequent analyses will be based on relative expression, so we won't round expression values or log-normalize.
```{r}
NanoString.mRNA.norm <- NanoStringNorm(
  x = NanoString.mRNA,
  CodeCount = "geo.mean",
  Background = "mean.2sd",
  SampleContent = "housekeeping.geo.mean",
  round.values = FALSE,
  take.log = FALSE)
```
Finally, we'll generate some plots to summarize the results of the normalization.
```{r}
pdf("./results/NanoStringNorm_results.pdf")
Plot.NanoStringNorm(
  x = NanoString.mRNA.norm,
  label.best.guess = TRUE,
  plot.type = c("cv", "mean.sd", "RNA.estimates", "missing",
                "norm.factors", "positive.controls"))
dev.off()
```

Export the normalized data for subsequent analyses in Matlab.
```{r}
write.table(x = NanoString.mRNA.norm$normalized.data, file = "./results/AD_NanoStringNorm_data.txt", sep = "\t")
```


