NanoStringNorm Final
========================================================

The following code is designed to import, process, and analyze NanoString data from Todd Golde's experiments in APP (Tg) and normal (WT) mice models, with and without over-expression of IL-10.

Load the required libraries.
```{r}
require("NanoStringNorm")
library("ggplot2")
```

We first import the data and create a data frame using `read.table`.
```{r}
NanoString.mRNA <- read.table("./data/AD_nanoString_data.txt", fill = TRUE, header = TRUE, 
                              as.is = TRUE)
```

Some of the gene have duplicate names, and names are less useful to us than the RefSeq accession numbers, so we'll replace the former with the latter.
```{r}
NanoString.mRNA$Name[1:240] <- NanoString.mRNA$Accession[1:240]
```

Next we define the housekeeping genes in the data.
```{r}
# NanoString.mRNA[NanoString.mRNA$Name %in% c("Cltc", "GAPDH", "Hprt1", "Pgk1", "Tubb5"), 
#                "Code.Class"] <- "Housekeeping"

NanoString.mRNA[NanoString.mRNA$Name %in% c("NM_001003908.1", "NM_001001303.1", "NM_013556.2", "NM_008828.2", "NM_011655.5"),
                "Code.Class"] <- "Housekeeping"
```

Here we extract sample subsets according to experiment.
```{r}
annoCols <- colnames(NanoString.mRNA)[c(1:3)]
sampleIDs <- colnames(NanoString.mRNA)[-c(1:3)]
tgSamples <- sampleIDs[c(1:12)]
wtSamples <- sampleIDs[c(13:21)]
oldSamples <- sampleIDs[c(22:27)]
```

We want to build a trait indicator vector to look at differential expression among Tg samples.
```{r}
NS.names <- names(NanoString.mRNA)[-c(1:3)]
tgExp <- (NS.names %in% tgSamples) + 1 # value 2 indicates current group
wtExp <- (NS.names %in% wtSamples) + 1
oldExp <- (NS.names %in% oldSamples) + 1

trait.experiment <- data.frame(
  row.names = NS.names,
  tgExp = tgExp,
  wtExp = wtExp,
  oldExp = oldExp)
```

```{r}
norm.comp.results.test <- norm.comp(
    x = NanoString.mRNA,
    CodeCount.methods = c("geo.mean"),
    Background.methods = c("mean.2sd"),
    SampleContent.methods = c("none", "housekeeping.sum", "housekeeping.geo.mean",
                              "total.sum", "low.cv.geo.mean", 
                              "top.mean", "top.geo.mean"),
    OtherNorm.methods = "none",
    histogram = FALSE,
    verbose = TRUE)
```

All of the above combinations of options yield virtually identical results for normalization. Given this observation, we'll go ahead and use the *recommended* set of options in the `NanoStringNorm` manual. Some of our subsequent analyses will be based on relative expression, so we won't round expression values or log-normalize.
```{r}
NanoString.mRNA.norm <- NanoStringNorm(
  x = NanoString.mRNA,
  CodeCount = "geo.mean",
  Background = "mean.2sd",
  SampleContent = "housekeeping.geo.mean",
  round.values = FALSE,
  take.log = TRUE,
  traits = trait.experiment)
```

Finally, we'll generate some plots to summarize the results of the normalization.
```{r}
pdf("./results/NanoStringNorm_results_hkLog.pdf")
Plot.NanoStringNorm(
  x = NanoString.mRNA.norm,
  label.best.guess = TRUE,
  plot.type = c("cv", "mean.sd", "RNA.estimates", "missing", "volcano",
                "norm.factors", "positive.controls"))
dev.off()
```

Export the normalized data for subsequent analyses in Matlab.
```{r}
write.table(x = NanoString.mRNA.norm$normalized.data, file = "./results/AD_NanoStringNorm_data_hkLog.txt", sep = "\t")
```


